\section{Deployment} % (fold)
\label{sec:deployment}

The installation of the whole \idx{ScaleNet} application is out of this project scope, however it is necessary to explain the upgrade from a previous version.
Since one of the goals was to be as less disruptive as possible, it should be fairly easy to deploy the new system into an existing installation.
The steps can be roughly summarized in three:

\begin{description}
  \item[Replace the \ida{PHP} sources.] The whole \texttt{scalenet} directory has to be copied over the old one in the same location.
  \item[Install the \ida{APE} server.] This just means installing a package.
  \item[Configure \idas{BIND}.] \ida{DNS} needs some custom rules.
\end{description}

\subsection{Install the APE Server} % (fold)
\label{sub:apeinstall}

The \ida{APE} download page \cite{ApeDownload} contains packages for different operating systems and architectures.
In this case, since the system is Debian-based we should use the DEB package.
Once the correct package is downloaded, it can be installed on the \idx{Application Server} by typing Listing~\vref{apeinstallation} from the same directory as the package is stored.

\begin{lstlisting}[label=apeinstallation,caption=APE installation command]
  sudo dpkg -i ape-1.0.i386.deb
\end{lstlisting}

After that, the \ida{APE} server daemon (\idc{aped}) is automatically started with the default configuration \cite{ApeSetup}.
It can be checked by visiting the url \url{webportal.imusu.mobile.dtrd.de:6969}.

% subsection apeinstall (end)

\subsection{Configure BIND} % (fold)
\label{sub:bind}

The \ida{IMS} core is the machine that provides the \idas{DNS} service through \idas{BIND}, and that service needs to be configured to allow the \ac{APE} server\index{APE server} to use a lot of different dynamic subdomains like \verb|1.ape.webportal|, \verb|2.ape.webportal|, \verb|567.ape.webportal|, etc.

This is how the \ac{APE} server\index{APE server} works by default, and it appears that there is no way to configure the \ac{APE} server\index{APE server} for using only one domain \cite{ApeConfig}.

So, in the file \verb|/etc/bind/imusu.dnszone| located in the \ida{IMS} core
we have to look for the \emph{webportal} entry and change that section to look
like Listing~\vref{bindconf}.

\begin{lstlisting}[label=bindconf,caption=BIND configuration]
  webportal               1D IN A         192.168.5.234
  ape.webportal           1D IN A         192.168.5.234
  *.ape.webportal         1D IN CNAME     ape.webportal
\end{lstlisting}

To apply the changes, we have to restart \idas{BIND} using the command in
Listing~\vref{bindrestart}.

\begin{lstlisting}[label=bindrestart,caption=BIND restart command]
  sudo /etc/init.d/bind restart
\end{lstlisting}
% subsection bind (end)
% section deployment (end)