\begin{acknowledgements}

I would like to thank my tutor José Ignacio for guiding me from Madrid, and Steffen, Dmitry and Hans for their warm welcome and constant help during my stay in Berlin.
Also thanks to T-Labs and the Erasmus Placement program for making this experience possible.

Thanks to Carolina, Javi, Laura, Lucas, Phillip and the rest of the friends that I made in Berlin, six months flew by because of their company.

Since this project culminates my studies in Universidad Carlos III, I also want to thank all the friends, colleagues and teachers that I met and that help me during all these years.
You made those endless hours working together in assignments seem bearable.

Finally, special thanks to my family for supporting me all these years and showing me all the patience in the world.

\begin{center}
  \line(1,0){250}
\end{center}

Me gustaría agradecer a mi tutor José Ignacio por orientarme desde Madrid, y a Steffen, Dmitry y Hans por su bienvenida y constante ayuda durante mi estancia en Berlín.
También agradecezo a los T-Labs y al programa Erasmus Placement el hacer posible esta experiencia.

Gracias a Carolina, Javi, Laura, Lucas, Phillip y al resto de amigos que hice en Berlín, por su compañía seis meses pasaron volando.

Dado que este proyecto culmina mis años de estudio en la Universidad Carlos III, también quiero agradecer a todos mis amigos, compañeros y profesores que he conocido y que me han ayudado durante todos estos años.
Habéis hecho fáciles de soportar las interminables horas trabajando juntos en prácticas.

Por último, un agradecimiento especial a mi familia por apoyarme todos estos años y por mostrarme toda la paciencia del mundo.

\end{acknowledgements}