\section{Motivation} % (fold)
\label{sec:motivation}

The current document describes the work done under a Erasmus Placement internship in the \ida{T-Labs} of Berlin, developed during six months.
This department is in charge of several research projects regarding, mainly, next generation networks.

The \idx{ScaleNet} project was an attempt to converge several networks and services under the same system.
These services may include voice and video telephony, mobile TV, video on demand, online gaming and internet access.
Network convergence is seen as the migration of heterogeneous physical and logical network elements of fixed and mobile networks into one single (IP based) infrastructure.

This project was effectively finished in 2009, and one of the outcomes is an \ida{IMS} demonstration environment.
Using that demonstrator, service applications built on top of \idx{ScaleNet} are shown to companies and organizations that are potentially interested in the technology.

One of those applications is the \ida{PNAI} application, a web based \ida{GUI} from where the user can control running sessions in their devices.
Sessions are simply instances of the previously mentioned services, and they could be from phone calls to video streaming.

The system can manage any number of devices for each user, so he can register his mobile phone, laptop, desktop or TV and he can manipulate the sessions running in any device.
It has also some social capabilities implemented, so the user can also control a list of friends and can \emph{buy}, start, transfer or duplicate sessions for any of his friends.

Additionally, another web application called \idx{IPTVplus} is available to the user.
The purpose of this application is the acquisition of video content: live streaming, video on demand, etc.
The user can buy any content and a new session should automatically start on his default device.

The project motivation was improving the user interface by redesigning this \ida{PNAI}, implementing some functional extensions and integrating the \idx{IPTVplus} directly into the \ida{PNAI}.
With a little rethinking, the \ida{GUI} could be modernized so that the demonstrator were stunningly effective.

Later in the project, this task extended to porting the same functionality to a mobile web application to modern devices such as the \idx{iPhone}, the \idx{iPad} or the \idx{Android} devices.
The system only supported \emph{legacy} mobile devices like PDAs controlled by stylus, that were not really so sexy even back in their day.
% section motivation (end)