\section{Document Structure} % (fold)
\label{sec:structure}

The work is mainly structured in four chapters, being this one the first.
In this chapter motivation and objectives for this project have been discussed, and the project phases have been delineated.

The second chapter explains the state of the art, that is, the existing technologies this project is based on.
This includes a somewhat detailed explanation of the \idx{ScaleNet} project in its current state, although only the relevant parts for this work have been reviewed.
Despite of the brevity of the existing documentation, some extra diagrams have been produced and added to help visualizing the system and more specifically the \ida{PNAI} interface.

Besides that system description, all the basic technologies used for the work are discussed, stating their utility and place in this work.
This sections talk about programming languages, platforms, libraries and techniques.

The third chapter concretes the actual development of this work, focusing on just the parts that have been modified.
New requirements are compiled and analyzed, producing use cases for all the new situations that need to be handled.
From that, the structure of the new system is described, starting from big components and going down to specific classes.

The development chapter continues with some implementation details that need further discussion or clarification.
The last section compiles the deployment instructions for upgrading a previous \idx{ScaleNet} installation to reckon the work in this project.

The four and last chapter evaluates this project, giving some conclusions to sum up the work made during those six months.
Also, it provides future insight and advice to possible directions expanding this project.

The end of this document collects an estimated budget, the bibliography used in this project, the definitions of the acronyms and an index of the key works.
% section structure (end)